Modelling the emergence of whisker barrels

### 2020-07-09 paper/briefpaper/brief.tex
Removes trailing whitespace on a couple of lines, otherwise no changes.
Latex tidying (line lengths)

\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig1.png}
    \caption{The emergence of whisker barrels. \textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1
      by \cite{zheng_signal_2001}, with black lines to delineate barrels and to
      measure departure (Honda-$\delta$; see \citealp{senft_mouse_1991}) from a
      perfect Voronoi tesselation. Right shows the initial distribution of axon
      branching density ($a$) for one thalamocortical projection, and two
      molecular guidance fields ($\rho$), \cmnt{where the domain $S$ has been
        traced from \textbf{A}}. \textbf{B} The strengths of interaction
      $\gamma$ with fields $\rho_1$ and $\rho_2$ are indicated for each of 41
      projections by the lengths of green and blue arrows respectively, assuming
      that similar fields aligned to the posterior-anterior and medial-lateral
      axes in the ventroposterior medial nucleus of the thalamus are sampled at
      the locations of putative barreloid centers (reconstructed from
      \citealp{haidarliu_size_2001}, their \MPone{Fig.\,5b}). \textbf{C} Results for the example simulation, with parameters
      \MPtwo{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
      $\epsilon=1.2$} and $\delta{t}=0.0001$. Colours indicate the thalamic
      projection for which the connection density is maximal and black lines
      delineate boundaries (see Movie
      S1). \textbf{D} Red dots show the \emph{Honda}--\metrics{$\delta(t)$} metric obtained from \metrics{the}
      simulation approaching that obtained from the \metrics{real} barrels in \textbf{A} (dotted
      line); black squares show the \emph{pattern difference} metric $\eta(t)$, and reveal the emergence of a correspondence between the real and simulated barrel shapes (units mm$^3$); \metrics{grey hexagons show \metrics{how selectively each cortical site is innervated};
        $\omega(t) = \oiint_{S} \mu(\mb{x},t) \mathrm{d}S$, where
        $\mu(\mb{x}) \equiv \mathrm{max}_i\big(c_i(\mb{x},t)\big)\big/\sum_{j=1}^{N} c_j(\mb{x},t)$.}
      \textbf{E} \metrics{Plotted across the cortical sheet, the selectivity
        develops to reveal an alignment with the emergent barrel boundary
        shapes}. \metrics{Greyscale colour indicates values of $\mu(\mb{x})$ and overlaid
        contours show $c_i > 0.95 \mathrm{max}_i\big(c_i(\mb{x})\big)$.} All scale bars 1\,mm.}
    \label{fig:main}
  \end{fullwidth}
\end{figure}

@@

% Full page width
\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig2.png}
    \caption{\MPthreePar{Exploring the parameter space. \textbf{A} Colour indicates the quality of the pattern at $t=30000$ steps, against three measures: Low values of \emph{Honda}--$\delta$ (top
        row) suggest a Voronoi-like pattern of fields. The \emph{pattern difference}, $\eta$ (middle row) measures the difference in the area and arrangement of barrels in real and simulated fields. The \emph{selectivity}, $\omega$ (bottom
        row), measures the specificity with which the cortical sheet is innervated. Colour maps are chosen so that lighter (orange and yellow) colours indicate higher quality patterns against each measure. White squares indicate combinations of parameters for which simulations were numerically unstable.
%
        The parameter space explored is three dimensional with the ratio $\alpha/\beta$ varying between plots, and the
        competition parameter $\epsilon$ and the diffusion constant $D$ varying within plots. An asterisk (*)
        marks the parameters used in Fig.\,\ref{fig:main}. Boxes i),
        ii)~and iii)~mark parameter sets for which corresponding patterns are
        shown in \textbf{B} (for $t=30000$ steps).
%
        \textbf{B} Varying the diffusion constant $D$ generates qualitatively different patterns. Higher values 
        cause expansions of the peripheral barrels and a corresponding compression of the inner barrels (i). Lower values instead cause an expansion of the central barrels and compression of the peripheral barrels (ii). Further reducing the rate of diffusion (iii), which is equivalent to increasing the size of the domain and hence simulating development in an animal with a larger cortex, causes a large area to be occupied by projections with intermediate interaction parameters; those with strong interaction parameters are compressed around the edge of the domain, and consequently a barrel pattern fails to form.}}
    \label{fig:paramsweep}
  \end{fullwidth}
\end{figure}

@@

% Full page width
\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig3.png}
    \caption{
      \MPthreeSens{Sensitivity analysis. \textbf{A} Metrics of map quality $\delta(t)$ (top row), $\omega(t)$ (middle row) and $\eta(t)$ (bottom row), were evaluated at $t=30000$ steps. The $y$-axes and colour scales have identical ranges
        to the colour scales in Fig.\,\ref{fig:paramsweep}, for easy comparison. \emph{Left column:} The effect of adding noise drawn from a uniform distribution,
        $(\gamma_{max}-\gamma_{min})\,U(0,\nu_\gamma)$,
        to the values of the interaction parameters, $\gamma$ used in
        Fig.\,\ref{fig:main}.
        %
        \emph{Middle column:} The effect of changing the magnitude and the length scale of
        noise applied to the guidance fields. Uniform random noise $(\rho_{max} - \rho_{min})\,U(0, \nu_{\rho})$ was added
        to each (hexagonal) element of $\rho_1(\mb{x})$ and $\rho_2(\mb{x})$ and the result
        was smoothed by convolution with a symmetric 2D
        Gaussian kernel of width $\sigma_\rho$. 
        %
        \emph{Right column:} The effect of setting the rotational angle
        of the linearly varying guidance field, $\rho_1(\mb{x})$, to $\phi_{\rho_1}$,
        and modifying its overall gain to $G_{\rho_1}$, whilst keeping the
        parameters of $\rho_2(\mb{x})$
        unchanged from those used in the example simulation, for which
        $\phi_{\rho_2}=84$\textdegree and $G_{\rho_2}=1$.
        %
        \textbf{B} Four ways in which the perturbations in \textbf{A} affect
        the patterns. i) The effect of significant interaction parameter noise
        results in topological defects. ii) High magnitude, short lengthscale noise in
        $\rho(\mb{x})$ leads to non-straight edges between adjacent barrels. iii)
 Reducing the slope of gradient $\rho_1$ (by a factor of 10)
causes barrel rows B, C and D to become `crushed'
        down the center line, and edge barrels to dominate. iv) Rotating $\rho_1$ by 20\textdegree~causes a slight distortion of the pattern, resulting in an overall anticlockwise rotation of the field locations.
      }
    }
    \label{fig:sens}
  \end{fullwidth}
\end{figure}

@@

\metrics{The \emph{pattern difference} metric, $\eta$, incorporates information about
  the differences in the areas of simulated and experimentally determined (real) barrels
  ($\mathcal{A}_i^{\mathrm{sim}}$ and $\mathcal{A}_i^{\mathrm{exp}}$), as well as
 information contained in the `adjacency vector' for each barrel, $\mathcal{V}_i$, the $j$-th element of which is the length of the border between barrel $i$ and $j$. For a well-formed barrel pattern, $\mathcal{V}_{i}$ is a sparse vector. A
  dimensionless quantity can be obtained from the scalar product of the simulated and experimental
  adjacency vectors:
  $\frac{1}{N} \sum_i \frac { \mathcal{V}_{i}^{\mathrm{sim}}}
  {b_{i}^{\mathrm{sim}} } \vcdot \frac{\mathcal{V}_{i}^{\mathrm{exp}}}{b_{i}^{\mathrm{exp}}}$, where e.g., 
  $b_i^{\mathrm{exp}}$, is the total length of the border around 
real barrel $i$. This quantity is small when the fields that form in simulation have dissimilar neighbour relations to those of the real barrels
  (e.g., Fig.\,\ref{fig:main}C, $t=1000$), and maximal for a precise topological map ($t=10000$).}
%
\metrics{A second comparison considers the mean magnitude of the difference between the simulated and experimental
  vectors: $\frac{1}{N} \sum_i \norm{ \mathcal{V}_{i}^{\mathrm{sim}} -
    \mathcal{V}_{i}^{\mathrm{exp}} } $. This tends to 0 for a perfect
  match and can separate patterns with straight boundaries from those with `noisy edges' (as in Fig.\,\ref{fig:sens}B\,ii).}
%
\metrics{We combined these terms into a single metric:}
%
\begin{equation} \label{eq:eta}
\color{colmetrics}
  \eta =
  \frac
  {
    \frac{1}{N} \sum_i \norm[\big]{ \mathcal{A}_i^{\mathrm{sim}} - \mathcal{A}_i^{\mathrm{exp}} }
    \times
    \frac{1}{N} \sum_i \norm[\big]{ \mathcal{V}_{i}^{\mathrm{sim}} -
      \mathcal{V}_{i}^{\mathrm{exp}} }
  }
  {
    \frac{1}{N}
    \sum_i \frac {\mathcal{V}_{i}^{\mathrm{sim}}} {b_{i}^{\mathrm{sim}}}
    \vcdot
    \frac {\mathcal{V}_{i}^{\mathrm{exp}}} {b_{i}^{\mathrm{exp}}}
  },
\color{black}
\end{equation}
%
\metrics{which has units of mm$^3$.}

>>>>

\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig1.png}
    \caption{The emergence of whisker barrels.
      %
      \textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1
      by \cite{zheng_signal_2001}, with black lines to delineate barrels and to
      measure departure (Honda-$\delta$; see \citealp{senft_mouse_1991}) from a
      perfect Voronoi tesselation. Right shows the initial distribution of axon
      branching density ($a$) for one thalamocortical projection, and two
      molecular guidance fields ($\rho$), \cmnt{where the domain $S$ has been
        traced from \textbf{A}}.
      %
      \textbf{B} The strengths of interaction $\gamma$ with fields $\rho_1$
      and $\rho_2$ are indicated for each of 41 projections by the lengths of
      green and blue arrows respectively, assuming that similar fields aligned
      to the posterior-anterior and medial-lateral axes in the ventroposterior
      medial nucleus of the thalamus are sampled at the locations of putative
      barreloid centers (reconstructed from \citealp{haidarliu_size_2001},
      their \MPone{Fig.\,5b}).
      %
      \textbf{C} Results for the example simulation, with parameters
      \MPtwo{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$,
        $\gamma\in\pm 2$, $\epsilon=1.2$} and $\delta{t}=0.0001$. Colours
      indicate the thalamic projection for which the connection density is
      maximal and black lines delineate boundaries (see Movie S1).
      %
      \textbf{D} Red dots show the \emph{Honda}--\metrics{$\delta(t)$} metric
      obtained from \metrics{the} simulation approaching that obtained from
      the \metrics{real} barrels in \textbf{A} (dotted line); black squares
      show the \emph{pattern difference} metric $\eta(t)$, and reveal the
      emergence of a correspondence between the real and simulated barrel
      shapes (units mm$^3$); \metrics{grey hexagons show how selectively each
        cortical site is innervated;
        $\omega(t) = \oiint_{S} \mu(\mb{x},t) \mathrm{d}S$, where
        $\mu(\mb{x}) \equiv \mathrm{max}_i\big(c_i(\mb{x},t)\big)\big/\sum_{j=1}^{N} c_j(\mb{x},t)$.}
      %
      \textbf{E} \metrics{Plotted across the cortical sheet, the selectivity
        develops to reveal an alignment with the emergent barrel boundary
        shapes}. \metrics{Greyscale colour indicates values of $\mu(\mb{x})$
        and overlaid contours show $c_i > 0.95
        \mathrm{max}_i\big(c_i(\mb{x})\big)$.} All scale bars 1\,mm.}
    \label{fig:main}
  \end{fullwidth}
\end{figure}

@@

% Full page width
\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig2.png}
    \caption{\MPthreePar{Exploring the parameter space.
        %
        \textbf{A} Colour indicates the quality of the pattern at $t=30000$
        steps, against three measures: Low values of \emph{Honda}--$\delta$
        (top row) suggest a Voronoi-like pattern of fields. The \emph{pattern
          difference}, $\eta$ (middle row) measures the difference in the area
        and arrangement of barrels in real and simulated fields. The
        \emph{selectivity}, $\omega$ (bottom row), measures the specificity
        with which the cortical sheet is innervated. Colour maps are chosen so
        that lighter (orange and yellow) colours indicate higher quality
        patterns against each measure. White squares indicate combinations of
        parameters for which simulations were numerically unstable.
        %
        The parameter space explored is three dimensional with the ratio
        $\alpha/\beta$ varying between plots, and the competition parameter
        $\epsilon$ and the diffusion constant $D$ varying within plots. An
        asterisk (*) marks the parameters used in Fig.\,\ref{fig:main}. Boxes
        i), ii)~and iii)~mark parameter sets for which corresponding patterns
        are shown in \textbf{B} (for $t=30000$ steps).
        %
        \textbf{B} Varying the diffusion constant $D$ generates qualitatively
        different patterns. Higher values cause expansions of the peripheral
        barrels and a corresponding compression of the inner barrels
        (i). Lower values instead cause an expansion of the central barrels
        and compression of the peripheral barrels (ii). Further reducing the
        rate of diffusion (iii), which is equivalent to increasing the size of
        the domain and hence simulating development in an animal with a larger
        cortex, causes a large area to be occupied by projections with
        intermediate interaction parameters; those with strong interaction
        parameters are compressed around the edge of the domain, and
        consequently a barrel pattern fails to form.}}
    \label{fig:paramsweep}
  \end{fullwidth}
\end{figure}

@@

% Full page width
\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig3.png}
    \caption{
      \MPthreeSens{Sensitivity analysis.
        %
        \textbf{A} Metrics of map quality $\delta(t)$ (top row), $\omega(t)$
        (middle row) and $\eta(t)$ (bottom row), were evaluated at $t=30000$
        steps. The $y$-axes and colour scales have identical ranges to the
        colour scales in Fig.\,\ref{fig:paramsweep}, for easy comparison.
        %
        \emph{Left column:} The effect of adding noise drawn from a uniform
        distribution, $(\gamma_{max}-\gamma_{min})\,U(0,\nu_\gamma)$, to the
        values of the interaction parameters, $\gamma$ used in
        Fig.\,\ref{fig:main}.
        %
        \emph{Middle column:} The effect of changing the magnitude and the
        length scale of noise applied to the guidance fields. Uniform random
        noise $(\rho_{max} - \rho_{min})\,U(0, \nu_{\rho})$ was added to each
        (hexagonal) element of $\rho_1(\mb{x})$ and $\rho_2(\mb{x})$ and the
        result was smoothed by convolution with a symmetric 2D Gaussian kernel
        of width $\sigma_\rho$.
        %
        \emph{Right column:} The effect of setting the rotational angle of the
        linearly varying guidance field, $\rho_1(\mb{x})$, to $\phi_{\rho_1}$,
        and modifying its overall gain to $G_{\rho_1}$, whilst keeping the
        parameters of $\rho_2(\mb{x})$ unchanged from those used in the
        example simulation, for which $\phi_{\rho_2}=84$\textdegree and
        $G_{\rho_2}=1$.
        %
        \textbf{B} Four ways in which the perturbations in \textbf{A} affect
        the patterns. i) The effect of significant interaction parameter noise
        results in topological defects. ii) High magnitude, short lengthscale
        noise in $\rho(\mb{x})$ leads to non-straight edges between adjacent
        barrels. iii) Reducing the slope of gradient $\rho_1$ (by a factor of
        10) causes barrel rows B, C and D to become `crushed' down the center
        line, and edge barrels to dominate. iv) Rotating $\rho_1$ by
        20\textdegree~causes a slight distortion of the pattern, resulting in
        an overall anticlockwise rotation of the field locations. }}
    \label{fig:sens}
  \end{fullwidth}
\end{figure}

@@

\metrics{The \emph{pattern difference} metric, $\eta$, incorporates
  information about the differences in the areas of simulated and
  experimentally determined (real) barrels ($\mathcal{A}_i^{\mathrm{sim}}$ and
  $\mathcal{A}_i^{\mathrm{exp}}$), as well as information contained in the
  `adjacency vector' for each barrel, $\mathcal{V}_i$, the $j$-th element of
  which is the length of the border between barrel $i$ and $j$. For a
  well-formed barrel pattern, $\mathcal{V}_{i}$ is a sparse vector. A
  dimensionless quantity can be obtained from the scalar product of the
  simulated and experimental adjacency vectors:
  %
  $\frac{1}{N} \sum_i \frac { \mathcal{V}_{i}^{\mathrm{sim}}}
  {b_{i}^{\mathrm{sim}} } \vcdot \frac{\mathcal{V}_{i}^{\mathrm{exp}}}{b_{i}^{\mathrm{exp}}}$,
  %
  where e.g., $b_i^{\mathrm{exp}}$, is the total length of the border around
  real barrel $i$. This quantity is small when the fields that form in
  simulation have dissimilar neighbour relations to those of the real barrels
  (e.g., Fig.\,\ref{fig:main}C, $t=1000$), and maximal for a precise
  topological map ($t=10000$).}
%
\metrics{A second comparison considers the mean magnitude of the difference
  between the simulated and experimental vectors: $\frac{1}{N} \sum_i \norm{
    \mathcal{V}_{i}^{\mathrm{sim}} - \mathcal{V}_{i}^{\mathrm{exp}} } $. This
  tends to 0 for a perfect match and can separate patterns with straight
  boundaries from those with `noisy edges' (as in Fig.\,\ref{fig:sens}B\,ii).}
%
\metrics{We combined these terms into a single metric:}
%
\begin{equation} \label{eq:eta}
\color{colmetrics}
  \eta =
  \frac
  {
    \frac{1}{N} \sum_i \norm[\big]{ \mathcal{A}_i^{\mathrm{sim}} - \mathcal{A}_i^{\mathrm{exp}} }
    \times
    \frac{1}{N} \sum_i \norm[\big]{ \mathcal{V}_{i}^{\mathrm{sim}} -
      \mathcal{V}_{i}^{\mathrm{exp}} }
  }
  {
    \frac{1}{N}
    \sum_i \frac {\mathcal{V}_{i}^{\mathrm{sim}}} {b_{i}^{\mathrm{sim}}}
    \vcdot
    \frac {\mathcal{V}_{i}^{\mathrm{exp}}} {b_{i}^{\mathrm{exp}}}
  },
\color{black}
\end{equation}
%
\metrics{which has units of mm$^3$.}
